Draft Report

### 2020-11-18 DraftReport.Rmd
Added model assumptions based on Model 3A.

Firstly, it is important to note that given each state (with the addition of the District of Columbia) is treated as a unique data point, the sample contains 51 data points. Although this size meets the general rule of an adequate sample size, as the analysis begins to factor in the wide range of potential population distributions and demographics within the states, it is clear that there is large variation within the sample. For example, the population density can have large variability depending on the population concentration among a few large areas as well as the level of uninhabited or sparsely inhabited land within a state, while the demographics of population can depend on a wide range of factors such as geographic location and employment opportunities. It is also important to note that these data points cannot be considered independent of each other, as there is a lot of interstate travel and regional similarities in terms of population and policy. This indicates that the sample cannot be assumed to be IID.

@@

Firstly, the most general requirement for the linear model is that data points be independent and identically distributed (IID). While for the most part the events of one state will primarily affect that single state, travel by individuals spreading the virus across state lines is an inevitable occurence. This may cause the spread of a virus within one state to be influenced by those around it, or generally by its position in the country. In addition in terms of demographics, all states are involved in the same market economy system even though there are economic performance divisions across the states. Moreover, in most states white race is still dominant in the population. Many states share the similar racial composition.

While this is acknowledged to occur, the effect is likely minor compared to the demographic and policy data collected state by state which would affect the spread of the virus. Therefore it is relatively safe for this assumption to be met. 
Secondly, the process being modeled must be described by a linear conditional expectation function of the variables which are included, in order for the linear regression to be valid. This can be assessed after the fact by visualizing the model over the range of each variable alongside the data, or also by visualizing the residuals over the range of each variable. This can be seen across the three models.

Thirdly, related to the relationship of the residuals to the data, homoscedasticity can be assessed by visualizing the size of residuals across the range of each variable. Alternatively, it can also be done via the Breusch-Pagan test to more objectively determine whether homoscedasticity is not present in the data.

Next, perfect colinearity is automatically detected by the R regression algorithms, which drop potential variables if they have complete colinearity with another variable. Near perfect colinearity is more problematic to detect, but a common hint to its existence is large standard errors on colinear features. Conceptually, the variables which may have an important colinearity in the final model are variables such as different percentage make ups of race and ethnicity, or related public policy. These variables are appropriately evaluatea and excluded throughout the model building process.

Finally, normally distributed errors can be easily assessed by a histogram of the residuals around zero.

>>>>

Firstly, it is important to note that given each state (with the addition of the District of Columbia) is treated as a unique data point, the sample contains 51 data points. Although this size meets the general rule of an adequate sample size, as the analysis begins to factor in the wide range of potential population distributions and demographics within the states, it is clear that there is large variation within the sample. For example, the population density can have large variability depending on the population concentration among a few large areas as well as the level of uninhabited or sparsely inhabited land within a state, while the demographics of population can depend on a wide range of factors such as geographic location and employment opportunities.

@@

Firstly, the most general requirement for the linear model is that data points be independent and identically distributed (IID). While for the most part the events of one state will primarily affect that single state, travel by individuals spreading the virus across state lines is an inevitable occurence. This may cause the spread of a virus within one state to be influenced by those around it, or generally by its position in the country. In addition, all states are involved in the same market economy system of the country to varying degrees, even though there are economic performance divisions across the states. Likewise, certain regions of the country may also have similar racial compositions due to a shared history across general geographic regions. While this is acknowledged to occur, the effect is likely minor compared to the policy data collected state by state which would affect the spread of the virus. Therefore it is relatively safe for this assumption to be met.

Secondly, the process being modeled must be described by a linear conditional expectation function of the variables which are included, in order for the linear regression to be valid. This can be assessed after the fact by visualizing the model over the range of each variable alongside the data, or also by visualizing the residuals over the range of each variable. For the purpose of this assumption and other model or data related assumptions, only the Model 3A will be used as it is the most inclusive. 
```{r}
avg_pop <- log10(mean(df$population_density))
avg_pov <- mean(df$poverty_pct)
avg_mask <- mean(df$mask_public_bool)
y_pop <- model3a$coefficients[1] + model3a$coefficients[2]*log10(df$population_density) + model3a$coefficients[3]*avg_pov + model3a$coefficients[4]*avg_mask
y_pov <- model3a$coefficients[1] + model3a$coefficients[2]*avg_pop + model3a$coefficients[3]*df$poverty_pct + model3a$coefficients[4]*avg_mask
y_pop_nomask <- model3a$coefficients[1] + model3a$coefficients[2]*avg_pop + model3a$coefficients[3]*avg_pov
y_pov_nomask <- model3a$coefficients[1] + model3a$coefficients[2]*avg_pop + model3a$coefficients[3]*df$poverty_pct

df %>% 
  ggplot() + 
  geom_point(aes(x = log10(population_density), y = case_rate_100k, group = factor(mask_public_bool), color = factor(mask_public_bool))) + 
  geom_line(aes(x = log10(population_density), y = y_pop, color = "red")) +
  geom_line(aes(x = log10(population_density), y = y_pop_nomask, color = "blue")) +
  scale_color_manual(labels = c("No", "Yes", "No", "Yes"), values = c("blue", "red", "blue", "red")) +
  labs(
    title = 'Linearity of Variables, Population Density', 
    x = 'Base 10 Log of Population Density', 
    y = 'Case Rate', 
    color = 'Mask Mandate'
  )
df %>% 
  ggplot() + 
  geom_point(aes(x = poverty_pct, y = case_rate_100k, group = factor(mask_public_bool), color = factor(mask_public_bool))) + 
  geom_line(aes(x = poverty_pct, y = y_pov, color = "red")) +
  geom_line(aes(x = poverty_pct, y = y_pov_nomask, color = "blue")) +
  scale_color_manual(labels = c("No", "Yes", "No", "Yes"), values = c("blue", "red", "blue", "red")) +
  labs(
    title = 'Linearity of Variables, Poverty Percentage', 
    x = 'Poverty Percentage', 
    y = 'Case Rate', 
    color = 'Mask Mandate'
  )
```
Linearity for the three regressed variables is demonstrated above. As can be seen when mask policy and population density are varied, the data generally follows a linear trend. This is similar when the poverty percentage is varied, although at higher levels of poverty the presence of a mask mandate made the case rate unexpectedly low. It appears that possible a negative quadratice function may fit the data better, although it appears close enough to linear that this assumption may hold.

Thirdly, related to the relationship of the residuals to the data, homoscedasticity can be assessed by visualizing the size of residuals across the range of each variable. Alternatively, it can also be done via the Breusch-Pagan test to more objectively determine whether homoscedasticity is not present in the data. Using the first method with the plots above, the residuals to each of the four regressions appear to be evenly spread throughout the range of the x axis. Therefore, this assumption can also be reasonably assumed to be met. The normal distribution of these errors also appears to be well met by the linear models, with most data points close to their respective line causing most residuals to be centered around zero.

Finally, perfect colinearity is automatically detected by the R regression algorithms, which drop potential variables if they have complete colinearity with another variable. Near perfect colinearity is more problematic to detect, but a common hint to its existence is large standard errors on colinear features. In Model 3A, all of the standard errors for model coefficients are at least less than half of the value of the coefficient itself. Therefore, near-perfect colinearity can be assumed to not be taking place between these variables. Conceptually, the variables which may have an important colinearity in the final model are variables such as different percentage make ups of race and ethnicity, or related public policy. These variables are appropriately evaluated and excluded throughout the model building process.
